\subsection{II.3 Proposed Improvements}

\paragraph{1. Add a budgeted status and recovery router.}
Route immediately after \texttt{execute\_step}: end without another LLM call when
the task is complete; review successful non-final steps; and send failures to a
\texttt{recover\_failure} node. That node should replan from the failed step, so
the fixed prefix ends at \texttt{step\_index-2}, using its execution and goal
verdicts as mode-aware feedback. Add \texttt{total\_steps} and
\texttt{adaptation\_count} to state, with context limits
\texttt{max\_total\_steps} and \texttt{max\_adaptations}; reject a tail longer
than the remaining budget and return an explicit exhausted verdict. This both
uses failure feedback and guarantees termination. The trade-off is more routing
logic and the risk of stopping a task whose useful recovery needed a larger
budget.

\paragraph{2. Review a typed, evidence-rich observation.}
Introduce an \texttt{AdaptiveObservation} containing all prior outcomes, the two
latest verdicts, accumulated summaries, workspace manifest, and compact artefact
checks such as file existence, schema and hashes. Use deterministic rules for
hard signals, including failure, missing dependencies and violated invariants;
reserve the LLM reviewer for ambiguous semantic changes. Require confidence and
evidence fields, and invoke a different critic or a second vote only below a
confidence threshold. This improves observability and separates system checks
from model judgement. It costs tokens and latency, while excessive raw context
could itself distract the reviewer, so artefacts should be summarised rather
than inserted wholesale.

\paragraph{3. Validate, persist and selectively roll back revisions.}
Insert \texttt{validate\_tail} before \texttt{merge\_plan}. It should enforce the
remaining step budget and agent constraints deterministically, then ask a plan
critic to check goal coverage, dependencies and consistency with completed
artefacts. A rejected candidate returns to \texttt{replan\_tail} within the same
adaptation budget. Extend the review schema with an optional
\texttt{restart\_from} index; when evidence invalidates prior work, archive the
affected artefacts and re-execute that suffix. After acceptance, atomically
checkpoint the old and new tails, reason, evidence and revision ID. This restores
verification, controlled rollback and a failure audit trail, at the cost of extra
critic calls, storage, and a new risk that an incorrect rollback discards valid
work.
